---
title: "Data_Preparation"
output: html_notebook
---

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(psych)
```


                      ################################# Reading in data and combining training and test data ########################################


```{r}
#read files
train=read.csv("./inputs/train.csv", stringsAsFactors=FALSE)
test=read.csv("./inputs/test.csv", stringsAsFactors=FALSE)

#check read in files
print(train)
print(test)
```

```{r}
#remove id from both train and test variables
train=train[,-1]
print(train)
test=test[,-1]
print(test)
```

```{r}
#combine train and test sets for analysis and cleaning
#test set has missing saleprice (uneqal dimension for merging with train set)
#make a copy of the test set before adding the extra column
test_tmp=test
test_tmp$SalePrice= NA
print(test_tmp)

#combine the added-column test set with the train set
all=rbind(train,test_tmp)
print(all)
```
```{r}
#finding the number of numeric codes
numericVars=which(sapply(all, is.numeric))
numericVarNames=names(numericVars)
sprintf("There are %s numeric variables", length(numericVars))
```



                    ######################################### Tabulate the number of missing values ####################################################


```{r}
#tabulate the number of missing values
miss_freq=sort(colSums(is.na(all)),decreasing=TRUE)
print(miss_freq)
sprintf("There are %s columns of missing values", sum(miss_freq>0))
```


                ######################################### To fix all 34 variables with missing values ################################################


1. PoolQC: Pool quality
		
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       NA	No Pool

```{r}
#replace missing values in PoolQC
#According to the data description, PoolQC has Excellent, Good, Average, Fair and No Pool levels. NA represents No Pool.
all$PoolQC[is.na(all$PoolQC)]='No Pool'
#convert the variable to ordinal class
values=c("No Pool"=0,"Po"=1,"Fa"=2,"TA"=3,"Gd"=4,"Ex"=5)
all$PoolQC=as.integer(revalue(all$PoolQC,values))
class(all$PoolQC)
```

```{r}
#assessing PoolQC with PoolArea
all[all$PoolQC==0 & all$PoolArea>0, c("PoolQC","PoolArea","OverallQual")]
```

```{r}
#this indicates some error in filling up of the missing values. '0' indicates no pool yet another variable showing its pool area.
#here we assume pools exist for these houses but without quality ratings.
#we will replace these values with reference to "OverallQual" variable
#there are 10 ordinal categories in "OverallQual" and we resampled it into 5 categories to match with PoolQC range.

all[2421,"PoolQC"]=2
all[2504,"PoolQC"]=3
all[2600,"PoolQC"]=2
```

2. MiscFeature: Miscellaneous feature not covered in other categories
		
       Elev	Elevator
       Gar2	2nd Garage (if not described in garage section)
       Othr	Other
       Shed	Shed (over 100 SF)
       TenC	Tennis Court
       NA	None

```{r}
#Replace missing values in MiscFeature
#MiscFeature account for the additional facilities that a house might possess, such as elevator, 2nd Gararge, Shed, Tennis Court and Others.
#The NAs are due to the house having no other additional facilities
#replacing 'NA' with 'None' and this feature should bein factor form.
all$MiscFeature[is.na(all$MiscFeature)]='None'
#convert the variable to factor class
all$MiscFeature=as.factor(all$MiscFeature)
class(all$MiscFeature)
```

3. Alley: Type of alley access to property

       Grvl	Gravel
       Pave	Paved
       NA 	No alley access

```{r}
#NA represent no alley access
all$Alley[is.na(all$Alley)]='No alley access'
#convert the variable to factor class
all$Alley=as.factor(all$Alley)
class(all$Alley)
```

4. Fence: Fence quality
		
       GdPrv	Good Privacy
       MnPrv	Minimum Privacy
       GdWo	Good Wood
       MnWw	Minimum Wood/Wire
       NA	No Fence

```{r}
#NA represents No Fence
all$Fence[is.na(all$Fence)]='No Fence'
#converting this to factor class
all$Fence=as.factor(all$Fence)
class(all$Fence)
```

5. FireplaceQu: Fireplace quality

       Ex	Excellent - Exceptional Masonry Fireplace
       Gd	Good - Masonry Fireplace in main level
       TA	Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement
       Fa	Fair - Prefabricated Fireplace in basement
       Po	Poor - Ben Franklin Stove
       NA	No Fireplace
       
```{r}
#replace NA with no Fireplace
all$FireplaceQu[is.na(all$FireplaceQu)]="No Fireplace"

#another variable that is closely related to FireplaceQu is Fireplaces
#and let's check whether the number of houses with no fireplace in "Fireplaces" equates with the count in "FireplaceQu"
sprintf("The number of No Fireplace in variables FireplaceQu and Fireplaces are equal: %s",sum(all$FireplaceQu=="No Fireplace") == sum(all$Fireplaces==0))

#convert the variable to ordinal class
values=c("No Fireplace"=0,"Po"=1,"Fa"=2,"TA"=3,"Gd"=4,"Ex"=5)
all$FireplaceQu=as.integer(revalue(all$FireplaceQu,values))
class(all$FireplaceQu)
```

6. LotFrontage: Linear feet of street connected to property

   Neighborhood: Physical locations within Ames city limits

       Blmngtn	Bloomington Heights
       Blueste	Bluestem
       BrDale	Briardale
       BrkSide	Brookside
       ClearCr	Clear Creek
       CollgCr	College Creek
       Crawfor	Crawford
       Edwards	Edwards
       Gilbert	Gilbert
       IDOTRR	Iowa DOT and Rail Road
       MeadowV	Meadow Village
       Mitchel	Mitchell
       Names	North Ames
       NoRidge	Northridge
       NPkVill	Northpark Villa
       NridgHt	Northridge Heights
       NWAmes	Northwest Ames
       OldTown	Old Town
       SWISU	South & West of Iowa State University
       Sawyer	Sawyer
       SawyerW	Sawyer West
       Somerst	Somerset
       StoneBr	Stone Brook
       Timber	Timberland
       Veenker	Veenker

```{r}
#get the index for the rows with missing values
Lot_NAs=all[is.na(all$LotFrontage),c("Neighborhood","LotFrontage")]
print(Lot_NAs)
idx=as.numeric(rownames(Lot_NAs))
print(class(idx))
print(length(idx))

# #we will use the mean distance of each "Neighbourhood" to replace the missing values in "LotFrontage"
for (i in idx){
      if(is.na(all$LotFrontage[i])){
              all[i,"LotFrontage"]=as.integer(mean(all$LotFrontage[all$Neighborhood==all[i,"Neighborhood"]], na.rm=TRUE))
        }
}
print(all[c(8,13,15,17),"LotFrontage"])
print(class(all$LotFrontage))
```

7. GarageYrBlt: Year garage was built

```{r}
#associate houses with missing values in this column as houses that did not rebuild or remodel its garage.
#replace this missing values with the year the house is built.
all[is.na(all$GarageYrBlt),"GarageYrBlt"]=all$YearBuilt[is.na(all$GarageYrBlt)]
class(all$GarageYrBlt)

#check whether the values have been replaced
sum(is.na(all$GarageYrBlt))
```
Putting all variables related to garage together:

Number of missing values in garag-related components:

GarageFinish: Interior finish of the garage = 159
GarageQual: Garage quality = 159
GarageCond: Garage condition = 159
GarageType: Garage location = 157
GarageCars: Size of garage in car capacity = 1
GarageArea: Size of garage in square feet = 1


```{r}
#GarageFinish, GarageQual, GarageCond, GarageType, GarageCars and GarageArea seems to be inter-related.
tmp=all[is.na(all$GarageFinish) & is.na(all$GarageQual) & is.na(all$GarageCond) & is.na(all$GarageType),]
sprintf("how many of these rows have NA values in GarageFinish, Qual, Cond, and Type: %s", dim(tmp)[1])
```

```{r}
#the above result indicates the the 157 missing values found in GarageType is also found in GarageFinish, Quala and Cond
#
all[is.na(all$GarageFinish) & is.na(all$GarageQual) & is.na(all$GarageCond) & !is.na(all$GarageType), c("GarageFinish","GarageQual","GarageCond","GarageType","GarageCars","GarageArea")]

#all[is.na(all$GarageCars) | is.na(all$GarageArea), c("GarageFinish","GarageQual","GarageCond","GarageType","GarageCars","GarageArea")]
```

8. GarageCars: Size of garage in car capacity 
9. GarageArea: Size of garage in square feet 

```{r}
# the record 2127 seems to have a garage as it has value for GarageCars and GarageArea
# on the other hand, 2577 does not seem to have a garage
# lets fix these two entries first

# function: finding mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

# entry 2127
all[2127,"GarageFinish"]=getmode(all$GarageFinish)
all[2127,"GarageQual"]=getmode(all$GarageQual)
all[2127,"GarageCond"]=getmode(all$GarageCond)

# entry 2577
all[2577,"GarageType"]=NA
all[2577,"GarageCars"]=0
all[2577,"GarageArea"]=0

#check the rectification
print(all[c(2127,2577),c("GarageFinish","GarageQual","GarageCond","GarageType","GarageCars","GarageArea")])
```
```{r}
tmp=all[is.na(all$GarageFinish) & is.na(all$GarageQual) & is.na(all$GarageCond) & is.na(all$GarageType),]
sprintf("how many of these rows have NA values in GarageFinish, Qual, Cond, and Type (after replacement): %s", dim(tmp)[1])
```
10. GarageFinish: Interior finish of the garage

       Fin	Finished
       RFn	Rough Finished	
       Unf	Unfinished
       NA	No Garage

```{r}
#the NAs indicate "No Garage"
#replace NAs with "No Garage"
all[is.na(all$GarageFinish),"GarageFinish"]="No Garage"

#check the number of NAs again
sum(is.na(all$GarageFinish))

# convert this variable to ordinal
values=c('No Garage'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)
all$GarageFinish=as.integer(revalue(all$GarageFinish, values))
class(all$GarageFinish)
```
11. GarageQual: Garage quality

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage

```{r}
#the NAs indicate "No Garage"
#replace NAs with "No Garage"
all[is.na(all$GarageQual),"GarageQual"]="No Garage"

#check the number of NAs again
sum(is.na(all$GarageQual))

#convert to ordinal
values=c("No Garage"=0,"Po"=1,"Fa"=2,"TA"=3,"Gd"=4,"Ex"=5)
all$GarageQual=as.integer(revalue(all$GarageQual, values))
class(all$GarageQual)
```
12. GarageCond: Garage condition

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage

```{r}
#the NAs indicate "No Garage"
#replace NAs with "No Garage"
all[is.na(all$GarageCond),"GarageCond"]="No Garage"

#convert to ordinal
values=c("No Garage"=0,"Po"=1,"Fa"=2,"TA"=3,"Gd"=4,"Ex"=5)
all$GarageCond=as.integer(revalue(all$GarageCond, values))
class(all$GarageCond)
```
13. GarageType: Garage location
		
       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage

```{r}
#the NAs indicate "No Garage"
#replace NAs with "No Garage"
all[is.na(all$GarageType),"GarageType"]="No Garage"

#convert to factor class
all$GarageType=as.factor(all$GarageType)
class(all$GarageType)
```

Putting all variables related to basement together:

No. of missing values in each variable:

BsmtCond: Evaluates the general condition of the basement = 82
BsmtExposure: Refers to walkout or garden level walls = 82  
BsmtQual: Evaluates the height of the basement = 81
BsmtFinType2: Rating of basement finished area = 80
BsmtFinType1: Rating of basement finished area = 79
BsmtFullBath: Basement full bathrooms = 2
BsmtHalfBath: Basement half bathrooms = 2
BsmtFinSF1: Type 1 finished square feet = 1
BsmtFinSF2: Type 2 finished square feet = 1
BsmtUnfSF: Unfinished square feet of basement area = 1
TotalBsmtSF: Total square feet of basement area = 1

```{r}
#check whether the 79 records with missing values in BsmtFinType1 correspond with those in BsmtCond, Exposure, Qual and FinType2
tmp=all[is.na(all$BsmtFinType1) & is.na(all$BsmtCond) & is.na(all$BsmtExposure) & is.na(all$BsmtQual) & is.na(all$BsmtFinType2),]
print(tmp)
sprintf("There are %s rows with NAs in bsmtfintype1, bsmtcond, bsmtexposure, bsmtqual and bsmtfintype2 concurrently ",dim(tmp)[1])

#finding the other records with NA in the basement variables.
all[!is.na(all$BsmtFinType1) & (is.na(all$BsmtCond)|is.na(all$BsmtExposure)|is.na(all$BsmtQual)|is.na(all$BsmtFinType2)),c("BsmtFinType1","BsmtCond","BsmtExposure","BsmtQual","BsmtFinType2")]
```

```{r}
#each row is observed to have a missing value out of five variables
#we will replace these NAs with the mode.

all[333,"BsmtFinType2"]=getmode(all$BsmtFinType2)
all[949,"BsmtExposure"]=getmode(all$BsmtExposure)
all[1488,"BsmtExposure"]=getmode(all$BsmtExposure)
all[2041,"BsmtCond"]=getmode(all$BsmtCond)
all[2186,"BsmtCond"]=getmode(all$BsmtCond)
all[2218,"BsmtQual"]=getmode(all$BsmtQual)
all[2219,"BsmtQual"]=getmode(all$BsmtQual)
all[2349,"BsmtExposure"]=getmode(all$BsmtExposure)
all[2525,"BsmtCond"]=getmode(all$BsmtCond)

all[c(333,949,1488,2041,2186,2218,2219,2349,2525),c("BsmtFinType1","BsmtCond","BsmtExposure","BsmtQual","BsmtFinType2")]
```
14. BsmtCond: Evaluates the general condition of the basement

       Ex	Excellent
       Gd	Good
       TA	Typical - slight dampness allowed
       Fa	Fair - dampness or some cracking or settling
       Po	Poor - Severe cracking, settling, or wetness
       NA	No Basement

```{r}
#the NAs indicate "No Basement"
#replace NAs with "No Basement"
all[is.na(all$BsmtCond),"BsmtCond"]="No Basement"

#convert to ordinal
values=c("No Basement"=0,"Po"=1,"Fa"=2,"TA"=3,"Gd"=4,"Ex"=5)
all$BsmtCond=as.integer(revalue(all$BsmtCond, values))
class(all$BsmtCond)
```
15. BsmtExposure: Refers to walkout or garden level walls

       Gd	Good Exposure
       Av	Average Exposure (split levels or foyers typically score average or above)	
       Mn	Mimimum Exposure
       No	No Exposure
       NA	No Basement
```{r}
#the NAs indicate "No Basement"
#replace NAs with "No Basement"
all[is.na(all$BsmtExposure),"BsmtExposure"]="No Basement"

#convert to ordinal
values=c("No Basement"=0,"No"=1,"Mn"=2,"Av"=3,"Gd"=4)
all$BsmtExposure=as.integer(revalue(all$BsmtExposure, values))
class(all$BsmtExposure)
```
16. BsmtQual: Evaluates the height of the basement

       Ex	Excellent (100+ inches)	
       Gd	Good (90-99 inches)
       TA	Typical (80-89 inches)
       Fa	Fair (70-79 inches)
       Po	Poor (<70 inches
       NA	No Basement
       
```{r}
#the NAs indicate "No Basement"
#replace NAs with "No Basement"
all[is.na(all$BsmtQual),"BsmtQual"]="No Basement"

#convert to ordinal
values=c("No Basement"=0,"Po"=1,"Fa"=2,"TA"=3,"Gd"=4,"Ex"=5)
all$BsmtQual=as.integer(revalue(all$BsmtQual, values))
class(all$BsmtQual)
```
17. BsmtFinType2: Rating of basement finished area (if multiple types)

       GLQ	Good Living Quarters
       ALQ	Average Living Quarters
       BLQ	Below Average Living Quarters	
       Rec	Average Rec Room
       LwQ	Low Quality
       Unf	Unfinshed
       NA	No Basement

```{r}
#the NAs indicate "No Basement"
#replace NAs with "No Basement"
all[is.na(all$BsmtFinType2),"BsmtFinType2"]="No Basement"

#convert to ordinal
values=c("No Basement"=0,"Unf"=1,"LwQ"=2,"Rec"=3,"BLQ"=4,"ALQ"=5,"GLQ"=6)
all$BsmtFinType2=as.integer(revalue(all$BsmtFinType2, values))
class(all$BsmtFinType2)
```
18. BsmtFinTyp1: Rating of basement finished area

       GLQ	Good Living Quarters
       ALQ	Average Living Quarters
       BLQ	Below Average Living Quarters	
       Rec	Average Rec Room
       LwQ	Low Quality
       Unf	Unfinshed
       NA	No Basement
       
```{r}
#the NAs indicate "No Basement"
#replace NAs with "No Basement"
all[is.na(all$BsmtFinType1),"BsmtFinType1"]="No Basement"

#convert to ordinal
values=c("No Basement"=0,"Unf"=1,"LwQ"=2,"Rec"=3,"BLQ"=4,"ALQ"=5,"GLQ"=6)
all$BsmtFinType1=as.integer(revalue(all$BsmtFinType1, values))
class(all$BsmtFinType1)
```
19. BsmtFullBath: Basement full bathrooms

```{r}
#replace NAs with zero
all[is.na(all$BsmtFullBath),"BsmtFullBath"]=0
hist(all$BsmtFullBath, main="Histogram of BsmtFullBath", xlab="number of full bathrooms in basement")
```
20. BsmtHalfBath: Basement half bathrooms

```{r}
#replace NAs with zero
all[is.na(all$BsmtHalfBath),"BsmtHalfBath"]=0
hist(all$BsmtHalfBath, main="Histogram of BsmtHalfBath", xlab="number of half bathrooms in basement")
```

21. BsmtFinSF1: Type 1 finished square feet
22.BsmtFinSF2: Type 2 finished square feet
23. BsmtUnfSF: Unfinished square feet of basement area
24. TotalBsmtSF: Total square feet of basement area

```{r}
#replace NAs with zero
#they are integers
all[is.na(all$BsmtFinSF1),"BsmtFinSF1"]=0
all[is.na(all$BsmtFinSF2),"BsmtFinSF2"]=0
all[is.na(all$BsmtUnfSF),"BsmtUnfSF"]=0
all[is.na(all$TotalBsmtSF),"TotalBsmtSF"]=0
```

Putting related variables, MasVnrType (24 missing values) and MasVnrArea(23 missing values), to assess.

```{r}
#the number of records with MasVnrType being NA given MasVnrArea is also NA.
tmp=all[is.na(all$MasVnrType) & is.na(all$MasVnrArea),]
sprintf("The number of records with with MasVnrType being NA given MasVnrArea is also NA: %s", dim(tmp)[1])

#finding the extra NA record of MasVnrType
all[is.na(all$MasVnrType) & !is.na(all$MasVnrArea), c("MasVnrType","MasVnrArea")]

#correct entry 2611
table(all$MasVnrType)
#"BrkFace" has the highest mode after "None"
all[2611,c("MasVnrType")]="BrkFace"
print(all[2611,c("MasVnrType","MasVnrArea")])

#replace NAs with None
all[is.na(all$MasVnrType),"MasVnrType"]="None"
```
25. MasVnrType: Masonry veneer type 

       BrkCmn	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       None	None
       Stone	Stone

```{r}
#check the relationship between the different category in MasVnrType and SalePrice
attach(all)
aggregate(all[,c("MasVnrType","SalePrice")],by=list(MasVnrType),FUN=mean,na.rm=TRUE)
#relationship seems to exist between "MasVnrType" and "SalePrice"
#convert to ordinal
values=c("BrkCmn"=0,"None"=1,"BrkFace"=2,"Stone"=3)
all$MasVnrType=as.integer(revalue(all$MasVnrType, values))
class(all$MasVnrType)
```
26. MasVnrArea: Masonry veneer area in square feet

```{r}
#replace NAs with zero
all[is.na(all$MasVnrArea),"MasVnrArea"]=0
```
27. MSZoning: Identifies the general zoning classification of the sale.
		
       A	Agriculture
       C	Commercial
       FV	Floating Village Residential
       I	Industrial
       RH	Residential High Density
       RL	Residential Low Density
       RP	Residential Low Density Park 
       RM	Residential Medium Density

```{r}
#input missing values with mode and convert to factor
all[is.na(all$MSZoning),"MSZoning"]=getmode(all$MSZoning)
all$MSZoning=as.factor(all$MSZoning)
class(all$MSZoning)
```
28. KitchenQual: Kitchen quality

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor

```{r}
#replace missing values with the mode
all[is.na(all$KitchenQual),"KitchenQual"]=getmode(all$KitchenQual)
sprintf("number of missing values in KitchenQual(after replacement):%s", sum(is.na(all$KitchenQual)))
#convert to ordinal
values=c("Po"=0,"Fa"=1,"TA"=2,"Gd"=3,"Ex"=4)
all$KitchenQual=as.integer(revalue(all$KitchenQual, values))
class(all$KitchenQual)
```
29. Utilities: Type of utilities available
		
       AllPub	All public Utilities (E,G,W,& S)	
       NoSewr	Electricity, Gas, and Water (Septic Tank)
       NoSeWa	Electricity and Gas Only
       ELO	Electricity only	

```{r}
#checking the frequency of each level
table(all$Utilities)

#the table shows that most of the rows are categorised as "AllPub"
#this variable hence serve little value in predicting the response varaiable
all$Utilities=NULL
```
30. Functional: Home functionality (Assume typical unless deductions are warranted)

       Typ	Typical Functionality
       Min1	Minor Deductions 1
       Min2	Minor Deductions 2
       Mod	Moderate Deductions
       Maj1	Major Deductions 1
       Maj2	Major Deductions 2
       Sev	Severely Damaged
       Sal	Salvage only

```{r}
#replace the missing value with the mode
all[is.na(all$Functional),"Functional"]=getmode(all$Functional)

#convert to ordinal
values=c("Sal"=0,"Sev"=1,"Maj2"=2,"Maj1"=3,"Mod"=4,"Min2"=5,"Min1"=6,"Typ"=7)
all$Functional=as.integer(revalue(all$Functional, values))
class(all$Functional)
table(all$Functional)

#checking the distribution of the Functional
table(all$Functional)
```
31. Exterior1st: Exterior covering on house

       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast	
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles

```{r}
#replace missing values with the mode
all[is.na(all$Exterior1st),"Exterior1st"]=getmode(all$Exterior1st)

#convert as factor
all$Exterior1st= as.factor(all$Exterior1st)
class(all$Exterior1st)
```
32. Exterior2nd: Exterior covering on house (if more than one material)

       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles

```{r}
#replace missing values with the mode
all[is.na(all$Exterior2nd),"Exterior2nd"]=getmode(all$Exterior2nd)

#convert as factor
all$Exterior2nd= as.factor(all$Exterior2nd)
class(all$Exterior2nd)
```
33.Electrical: Electrical system

       SBrkr	Standard Circuit Breakers & Romex
       FuseA	Fuse Box over 60 AMP and all Romex wiring (Average)	
       FuseF	60 AMP Fuse Box and mostly Romex wiring (Fair)
       FuseP	60 AMP Fuse Box and mostly knob & tube wiring (poor)
       Mix	Mixed

```{r}
#replace missing value with mode and convert variable to factor
all[is.na(all$Electrical),"Electrical"]=getmode(all$Electrical)

all$Electrical=as.factor(all$Electrical)
class(all$Electrical)
table(all$Electrical)
```
34. SaleType: Type of sale
		
       WD 	Warranty Deed - Conventional
       CWD	Warranty Deed - Cash
       VWD	Warranty Deed - VA Loan
       New	Home just constructed and sold
       COD	Court Officer Deed/Estate
       Con	Contract 15% Down payment regular terms
       ConLw	Contract Low Down payment and low interest
       ConLI	Contract Low Interest
       ConLD	Contract Low Down
       Oth	Other

```{r}
#replace missing value with mode and convert variable to factor
all[is.na(all$SaleType),"SaleType"]=getmode(all$SaleType)

all$SaleType=as.factor(all$SaleType)
class(all$SaleType)
table(all$SaleType)
```

                  ################################### Convert the remaining variables to appropriate class #######################################


a. Foundation: Type of foundation
		
       BrkTil	Brick & Tile
       CBlock	Cinder Block
       PConc	Poured Contrete	
       Slab	Slab
       Stone	Stone
       Wood	Wood


```{r}
#converting into factor class
all$Foundation=as.factor(all$Foundation)
class(all$Foundation)
```
b. Heating: Type of heating
		
       Floor	Floor Furnace
       GasA	Gas forced warm air furnace
       GasW	Gas hot water or steam heat
       Grav	Gravity furnace	
       OthW	Hot water or steam heat other than gas
       Wall	Wall furnace

```{r}
#converting into factor class
all$Heating=as.factor(all$Heating)
class(all$Heating)
```
c. HeatingQC: Heating quality and condition

       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor

```{r}
#converting into ordinal
values=c("Po"=0,"Fa"=1,"TA"=2,"Gd"=3,"Ex"=4)
all$HeatingQC=as.integer(revalue(all$HeatingQC, values))
class(all$HeatingQC)
```
d. CentralAir: Central air conditioning

       N	No
       Y	Yes

```{r}
#convert to ordinal
values=c("N"=0,"Y"=1)
all$CentralAir=as.integer(revalue(all$CentralAir, values))
class(all$CentralAir)
```
e. RoofStyle: Type of roof

       Flat	Flat
       Gable	Gable
       Gambrel	Gabrel (Barn)
       Hip	Hip
       Mansard	Mansard
       Shed	Shed

```{r}
#converting into factor class
all$RoofStyle=as.factor(all$RoofStyle)
class(all$RoofStyle)
```
f.RoofMatl: Roof material

       ClyTile	Clay or Tile
       CompShg	Standard (Composite) Shingle
       Membran	Membrane
       Metal	Metal
       Roll	Roll
       Tar&Grv	Gravel & Tar
       WdShake	Wood Shakes
       WdShngl	Wood Shingles

```{r}
#converting into factor class
all$RoofMatl=as.factor(all$RoofMatl)
class(all$RoofMatl)
```
g. LandContour: Flatness of the property

       Lvl	Near Flat/Level	
       Bnk	Banked - Quick and significant rise from street grade to building
       HLS	Hillside - Significant slope from side to side
       Low	Depression

```{r}
#converting into factor class
all$LandContour=as.factor(all$LandContour)
class(all$LandContour)
```
h.LandSlope: Slope of property
		
       Gtl	Gentle slope
       Mod	Moderate Slope	
       Sev	Severe Slope

```{r}
#converting to ordinal
values=c("Sev"=0,"Mod"=1,"Gtl"=2)
all$LandSlope=as.integer(revalue(all$LandSlope, values))
class(all$LandSlope)
```
i. BldgType: Type of dwelling
		
       1Fam	Single-family Detached	
       2FmCon	Two-family Conversion; originally built as one-family dwelling
       Duplx	Duplex
       TwnhsE	Townhouse End Unit
       TwnhsI	Townhouse Inside Unit
       
```{r}
#convert to factor
all$BldgType=as.factor(all$BldgType)
class(all$BldgType)
```
j. HouseStyle: Style of dwelling
	
       1Story	One story
       1.5Fin	One and one-half story: 2nd level finished
       1.5Unf	One and one-half story: 2nd level unfinished
       2Story	Two story
       2.5Fin	Two and one-half story: 2nd level finished
       2.5Unf	Two and one-half story: 2nd level unfinished
       SFoyer	Split Foyer
       SLvl	Split Level

```{r}
#convert to factor
all$HouseStyle=as.factor(all$HouseStyle)
class(all$HouseStyle)
```
k. Neighborhood: Physical locations within Ames city limits

       Blmngtn	Bloomington Heights
       Blueste	Bluestem
       BrDale	Briardale
       BrkSide	Brookside
       ClearCr	Clear Creek
       CollgCr	College Creek
       Crawfor	Crawford
       Edwards	Edwards
       Gilbert	Gilbert
       IDOTRR	Iowa DOT and Rail Road
       MeadowV	Meadow Village
       Mitchel	Mitchell
       Names	North Ames
       NoRidge	Northridge
       NPkVill	Northpark Villa
       NridgHt	Northridge Heights
       NWAmes	Northwest Ames
       OldTown	Old Town
       SWISU	South & West of Iowa State University
       Sawyer	Sawyer
       SawyerW	Sawyer West
       Somerst	Somerset
       StoneBr	Stone Brook
       Timber	Timberland
       Veenker	Veenker

```{r}
#convert to factor
all$Neighborhood=as.factor(all$Neighborhood)
class(all$Neighborhood)
```
l. Condition1: Proximity to various conditions
	
       Artery	Adjacent to arterial street
       Feedr	Adjacent to feeder street	
       Norm	Normal	
       RRNn	Within 200' of North-South Railroad
       RRAn	Adjacent to North-South Railroad
       PosN	Near positive off-site feature--park, greenbelt, etc.
       PosA	Adjacent to postive off-site feature
       RRNe	Within 200' of East-West Railroad
       RRAe	Adjacent to East-West Railroad

```{r}
#convert to factor
all$Condition1=as.factor(all$Condition1)
class(all$Condition1)
```
m. Condition2: Proximity to various conditions (if more than one is present)
		
       Artery	Adjacent to arterial street
       Feedr	Adjacent to feeder street	
       Norm	Normal	
       RRNn	Within 200' of North-South Railroad
       RRAn	Adjacent to North-South Railroad
       PosN	Near positive off-site feature--park, greenbelt, etc.
       PosA	Adjacent to postive off-site feature
       RRNe	Within 200' of East-West Railroad
       RRAe	Adjacent to East-West Railroad

```{r}
#convert to factor
all$Condition2=as.factor(all$Condition2)
class(all$Condition2)
```
n. Street: Type of road access to property

       Grvl	Gravel	
       Pave	Paved
       
```{r}
#convert to ordinal
values=c("Grvl"=0,"Pave"=1)
all$Street=as.integer(revalue(all$Street, values))
class(all$Street)
```
o. PavedDrive: Paved driveway

       Y	Paved 
       P	Partial Pavement
       N	Dirt/Gravel
       
```{r}
#convert to ordinal
values=c("N"=0,"P"=1,"Y"=2)
all$PavedDrive=as.integer(revalue(all$PavedDrive, values))
class(all$PavedDrive)
```
p. ExterQual: Evaluates the quality of the material on the exterior 
		
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor
       
```{r}
#convert to ordinal
values=c("Po"=0,"Fa"=1,"TA"=2,"Gd"=3,"Ex"=4)
all$ExterQual=as.integer(revalue(all$ExterQual, values))
class(all$ExterQual)
#check the distribution of the ordinal variable
table(all$ExterQual)
```
q. ExterCond: Evaluates the present condition of the material on the exterior
		
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       Po	Poor
       
```{r}
#convert to ordinal
values=c("Po"=0,"Fa"=1,"TA"=2,"Gd"=3,"Ex"=4)
all$ExterCond=as.integer(revalue(all$ExterCond, values))
class(all$ExterCond)
#check the distribution of the ordinal variable
table(all$ExterCond)
```
r. YrSold: Year Sold (YYYY)

s. MoSold: Month Sold (MM)

```{r}
#the prices of houses should be cheaper in earlier years.
#convert YrSold to ordinal
sprintf("Is data type of YrSold integer: %s", is.integer(YrSold))
#convert to factor
sprintf("Is data type of MoSold integer: %s", is.integer(MoSold))
all$MoSold=as.factor(all$MoSold)
class(all$MoSold)
```
t. MSSubClass: Identifies the type of dwelling involved in the sale.	

        20	1-STORY 1946 & NEWER ALL STYLES
        30	1-STORY 1945 & OLDER
        40	1-STORY W/FINISHED ATTIC ALL AGES
        45	1-1/2 STORY - UNFINISHED ALL AGES
        50	1-1/2 STORY FINISHED ALL AGES
        60	2-STORY 1946 & NEWER
        70	2-STORY 1945 & OLDER
        75	2-1/2 STORY ALL AGES
        80	SPLIT OR MULTI-LEVEL
        85	SPLIT FOYER
        90	DUPLEX - ALL STYLES AND AGES
       120	1-STORY PUD (Planned Unit Development) - 1946 & NEWER
       150	1-1/2 STORY PUD - ALL AGES
       160	2-STORY PUD - 1946 & NEWER
       180	PUD - MULTILEVEL - INCL SPLIT LEV/FOYER
       190	2 FAMILY CONVERSION - ALL STYLES AND AGES

```{r}
#convert to factor
all$MSSubClass=as.factor(all$MSSubClass)
str(all$MSSubClass)
```
u. LotShape: General shape of property

       Reg	Regular	
       IR1	Slightly irregular
       IR2	Moderately Irregular
       IR3	Irregular

```{r}
#convert LotShape as factor
values=c('IR3'=0, 'IR2'=1, 'IR1'=2, 'Reg'=3)
all$LotShape=as.integer(revalue(all$LotShape,values))
class(all$LotShape)
#check the distribution of the data
table(all$LotShape)
```
v. LotConfig: Lot configuration

       Inside	Inside lot
       Corner	Corner lot
       CulDSac	Cul-de-sac
       FR2	Frontage on 2 sides of property
       FR3	Frontage on 3 sides of property

```{r}
#convert LotConfig as factor
all$LotConfig=as.factor(all$LotConfig)
class(all$LotConfig)
```

w.SaleCondition: Condition of sale

       Normal	Normal Sale
       Abnorml	Abnormal Sale -  trade, foreclosure, short sale
       AdjLand	Adjoining Land Purchase
       Alloca	Allocation - two linked properties with separate deeds, typically condo with a garage unit	
       Family	Sale between family members
       Partial	Home was not completed when last assessed (associated with New Homes)

```{r}
#convert as factor
all$SaleCondition=as.factor(all$SaleCondition)
class(all$SaleCondition)
```


                    ################################# Check the number of missing data and structure of the data  #####################################


```{r}
#tabulate the number of missing values
miss_freq=sort(colSums(is.na(all)),decreasing=TRUE)
print(miss_freq)
sprintf("There are %s columns of missing values", sum(miss_freq>0))

#check the structure of the data
str(all)
```

```{r}
#save dataframe with its formatting.
saveRDS(all,file="all_DP.rds")
```

